ms_8

### 2020-04-24 ms_8.Rmd
created graphs for obesity data, working on food access data

food_access <- read_excel("Food Accessibility.xlsx", sheet = 3)

@@

obesity <- obesity_x %>% 
    rename("state" = LocationAbbr,
           "year" = YearStart,
           "bmi" = Data_Value,
           "conf.low" = Low_Confidence_Limit,
           "conf.high" = High_Confidence_Limit,
           "sample_size" = Sample_Size,
           "age" = "Age(years)",
           "education" = Education,
           "gender" = Gender,
           "class" = Class,
           "race" = "Race/Ethnicity",
           "income" = Income) %>% 
    filter(class == "Obesity / Weight Status",
           !is.na(bmi)) %>% 
    select(-class)

obesity_mean <- obesity %>% 
    group_by(state, year) %>% 
    summarise(mean_bmi = mean(bmi)) %>% 
    filter(state != "US")

plot_usmap(data = obesity_mean, values = "mean_bmi", color = "red", labels = FALSE) + 
    scale_fill_continuous(name = "bmi") + 
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5)) + 
    scale_fill_continuous(low = "white", high = "red", name = "BMI") +
    labs(title = "BMI in the US from 2011-2018",
         subtitle = "BMI ",
         caption = "Data from Center for Disease Control") +
    theme_void() +
    transition_time(year)

View(obesity_mean)

>>>>

food_access_x <- read_excel("Food Accessibility.xlsx", sheet = 3)

@@

obesity <- obesity_x %>% 
    rename("state" = LocationAbbr,
           "year" = YearStart,
           "perc_obese" = Data_Value,
           "conf.low" = Low_Confidence_Limit,
           "conf.high" = High_Confidence_Limit,
           "sample_size" = Sample_Size,
           "age" = "Age(years)",
           "education" = Education,
           "gender" = Gender,
           "class" = Class,
           "race" = "Race/Ethnicity",
           "income" = Income) %>% 
    filter(class == "Obesity / Weight Status",
           !is.na(perc_obese)) %>% 
    select(-class)

```{r obesity mean plot, include=FALSE}

obesity_mean <- obesity %>% 
    group_by(state, year) %>% 
    summarise(mean_perc = mean(perc_obese)) %>% 
    filter(state != "US")

obesity_mean_plot <- plot_usmap(data = obesity_mean, values = "mean_perc", color = "red", labels = FALSE) +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5)) + 
    scale_fill_continuous(low = "white", high = "red", name = "Percent Obese") +
    labs(title = "Percent of Population (age 18+) Classified as Obese or Overweight in the US \n Year: {as.integer(frame_time)}",
         subtitle = "Percent of population tends to increase as time goes on",
         caption = "Data from Center for Disease Control") +
    theme_void() +
    transition_time(year)

animate(obesity_mean_plot)

anim_save("plots/obesity_mean_usmap.gif", plot = last_animation())

```{r obesity by region}

obesity_region <- obesity_mean %>% 
    mutate(region = case_when(
        state %in% c("CT", "MA", "ME", "NH", "RI", "VT", "NJ", "NY", "PA") ~ "Northeast",
        state %in% c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD") ~ "Midwest",
        state %in% c("DC", "DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", 
                     "LA", "OK", "TX") ~ "South",
        state %in% c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA") ~ "West")
    ) %>% 
    mutate(year = as.numeric(year)) %>% 
  mutate(mean_perc = round(mean_perc, digits = 0))

obesity_region_plot <- obesity_region %>% 
    filter(!is.na(region)) %>%
    ggplot(aes(color = region, x = year, y = mean_perc)) +
    geom_jitter(aes(alpha = 0.1), show.legend = FALSE) +
    theme_minimal() +
    labs(title = "Percent of Population Classified as Obese or Overweight \n in US Over Time by Region",
         caption = "Data from Center for Disease Control",
         x = "Mean Percent of Population",
         y = "Year",
         color = "Region") +
    geom_smooth(method = "lm", se = FALSE, formula = y ~ x) +
    facet_wrap(~ region)
obesity_region_plot

ggsave("plots/obesity_region_plot.png", plot = last_plot())

```{r clean food access data, include=FALSE}

food_access_x <- food_access_x %>% 
    select(State, County, Urban, POP2010, LILATracts_1And10, PovertyRate, MedianFamilyIncome, LA1and10, LAPOP1_10)

food_access <- food_access_x %>% 
    rename("state" = State,
           "county" = County,
           "popul" = POP2010,
           "urban" = Urban,
           "lia_1_10" = LILATracts_1And10,
           "poverty_rate" = PovertyRate,
           "med_income" = MedianFamilyIncome,
           "la_1_10" = LA1and10,
           "la_popul" = LAPOP1_10) %>% 
    mutate(urban = as.factor(urban),
           lia_1_10 = as.factor(lia_1_10),
           la_1_10 = as.factor(la_1_10))

food_access$state = state.abb[match(food_access$state, state.name)]

View(food_access)
